---
title: "Series3"
author: "Group E"
date: "February 18, 2016"
output: pdf_document
---
##Preliminary Data Prep
Load and prepare the data for analysis. This includes taking 5 observations out of the data to be used to test the model accuracy. This hold-out series is called TS3No5.
```{r, echo=FALSE,include=F}

library(xlsx)
library(e1071)
library(dplyr)
library(forecast)

setwd("/home/ab/Documents/MBD/time_series/TS_session10_ex/data")
TS3= read.xlsx("Series3.xls",1)
TS3No5= as.data.frame(TS3[1:92,1:2])
```

##Step 1 - Visualise the Time Series
In this step, we check if the data is stationary in the mean and varience. This helps us determine what transformation is required.
We notice that the series is not stationary in the mean or the varience. In fact, there appears to be seasonaility, ACF and PACF showing a significant spike at every twelth value. 

```{r,echo=F}
X=TS3No5$SALES
tsdisplay(X, main = "Monthly Clothes Sales",lag.max = 36)
```

Skewness & Kurtosis are test to help determine if the data is normal or not. 
In normally distributed data, Skewness is 0 Kurtosis =is.
```{r, echo=F}
skewness(X, na.rm=TRUE)
kurtosis(X, na.rm=TRUE)
```

##Step 2 - Model
Given that the series is not stationary in the mean and there is seasonaility, we apply a model that accounts for both seasonaility and trend. The model below has coefficients different from 0.
```{r}
mod3 = arima(X, order=c(0,0,0), seasonal=list(order=c(1,0,0), period=12))
mod3
```

##Step 3 - Check Residuals
After we plot the residuals, we see the mode is still not stationary. We have missed a pattern in the data that helps explain the series.
```{r}
tsdisplay(mod3$residuals)
```

Using the residual plot from the first model, we fit the model below.
```{r}
mod3_1 =  arima(X,order = c(2, 1, 1), seasonal = list(order = c(2, 0, 1), period = 12))
```

##Step 4: Check if residuals are white noise
Its residuals appear to be white noise.
```{r}
tsdisplay(mod3_1$residuals)
```

The P-Value above .05 in the Ljung Box test indicates there is no structure in the residuals.
```{r}
Box.test(mod3_1$residuals,lag = 36,type = 'Ljung')
```

The shapiro-test P-Value of greater than 0.05 indicates the residuals are normal. 
```{r}
shapiro.test(mod3_1$residuals)
```

##Step 5: Forecase
```{r}
plot(forecast(mod3_1))
```


