---
title: "Series1"
author: "Group E"
date: "February 18, 2016"
output: pdf_document
---

```{r,echo=F,include=FALSE}
library(xlsx)
library(forecast)
library(tseries)
```

##Preliminary Data Prep
Load and prepare the data for analysis.
```{r,echo=F,include=FALSE}
setwd("/home/ab/Documents/MBD/time_series/TS_session10_ex/data")
ts1 <- read.xlsx("Series1.xls",1)
ts1 <- data.frame(ts1)[-1,]
ts1.a <- ts1[,2][1:565] # returns
ts1.b <- ts1[,3][1:565] #Moving Averages
ts1.a = na.omit(ts1.a)
ts1.b = na.omit(ts1.b)
```

##Step 1 - Visualise the Time Series and run an ACF and PACF.
In this step, we check if the data is stationary in the mean and varience. This helps us determine what transformation might be required.
```{r}
tsdisplay(ts1.a, main = "Time Series, ACF & PCF for TS1")
```

```{r}
tsdisplay(ts1.b, main = "Time Series, ACF & PCF for TS1 - MA")
```

##Step 2 - Analyze time series
For Microsoft returns, we have what appears to be white noise: The mean is 0, the variance is constant and the ACF and PACF are both 0. This means that we cannot forecast with this time series. We can explore this a little further with the Box Test and Shapiro test to see if we have strict white noise or Guassian white noise.  
```{r}
ts1.a.bt = Box.test(ts1.a,lag=10) 
ts1.a.bt$p.value # p-value is large, we do not reject the null of independence
ts1.a.st = shapiro.test(ts1.a) 
ts1.a.st$p.value # p-value is small, we reject the null of normality. 

# It appears we have strict white noise but not Guassian white noise. 
#Nontheless, we cannot create a valid model with white noise. We will leave this one here. 

# Turning now to time series B (smoothed MA=3) we see from the graphs that it is already stationary. 
#We can verify this with the Dickey Fuller Test. 

ts1.b.adf = adf.test(ts1.b)
ts1.b.adf$p.value # The p_value is small, we reject the null hypothesis that the series is non-stationary. 

# Lets look at some models for time series B since we know it is stationary.
# First we notice that our ACF and PACF graphs have lag cut-offs at ~3 and 1 respectively. 
#Therefore lets try a (1,0,3) model. 
b.fit <- arima(ts1.b, order = c(1,0,3)) 
b.fit$aic # aic is low 
```

##Step 3 - Analyze Residuals
```{r,echo=F,}
# Lets look at the ACF and PACF of the residuals
tsdisplay(b.fit$residuals)
# Residuals appear to be white noise
```

Box Test and Shapiro Test for Residuals to check for independence and normality of residuals
```{r,echo=F,}
Box.test(b.fit$residuals, lag = 10) # p-value is high, cannot reject null of independence
shapiro.test(b.fit$residuals) # p-value is small therefore we reject the null hypothesis of normality.
# We have strict white noise but not guassian white noise
```

##Step 4 - Forecast
Now we can make our predictions for the next 6 steps ahead 
```{r,echo=F,fig.height=3 }
ts1b.pred <- predict(b.fit, n.ahead = 6)
as(ts1b.pred,"list")
ts.plot(ts1.b)
lines(ts1b.pred$pred+1.96*ts1b.pred$se,col="red",lty=3)
lines(ts1b.pred$pred-1.96*ts1b.pred$se,col="red",lty=3)

```
